Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants\thanks{Supported by EPSRC grant EP/J012564/1, ERC project 280053 (CPROVER) and the H2020 FET OPEN 712689 SC$^2$.}

### 2017-04-28 cav2017/cav2017.tex
editing

Linear Time Invariant (LTI) models represent a broad class of dynamical
systems with significant impact in numerous application areas as in the life
sciences, robotics, and engineering~\cite{astrom1997computer,Franklin15}. 
Synthesis of controllers for LTI models is broadly investigated, however the
use of digital control architectures adds new challenges due to the effects
of finite-precision arithmetic, time discretization, and quantization noise,
which are typically introduced by Analogue-to-Digital (ADC) and
Digital-to-Analogue (DAC) conversion.  While research on digital control is
well developed~\cite{astrom1997computer}, automated and sound control
synthesis is challenging, particularly when the synthesis objective goes
beyond classical stability.  There are recent methods for verifying more
general reachability properties of a given controller~\cite{FLD+11,Fre05}. 
However, these methods have not been generalized to control synthesis. 
We~remark that a synthesis algorithm that guarantees satability does not
ensure safety: the system might transitively visit an unsafe state that
results in an unrecoverable failure.

@@

We give two alternative approaches for synthesizing digital controllers for
state-space physical plants, both based on CounterExample Guided Inductive
Synthesis (CEGIS)~\cite{sketch}.  We prove their soundness by quantifying
errors caused digitalization and quantization effects that arise when
time/value-discrete digital controller interacts with the continuous plant.

@@

We provide experimental results showing that both approaches are able to
efficiently synthesize stable and safe controllers for a set of intricate
physical plant models taken from the digital control literature: the median
run-time for our benchmark set is $7.9$ seconds, and most controllers can be
synthesized in less than $17.2$ seconds.  We further show that, in a direct
comparison, the abstraction-based approach (second) lowers the median
run-time of our benchmarks by a factor of seven over the first approach
based on the unfolding of the dynamics.

@@

\begin{enumerate}
%
\item  We compute state-feedback controllers that are provably stable and in
  addition guaranteee a given safety property.  Existing methods for
  controller synthesis rely on transfer function representations, which are
  insufficient to prove safety requirements.
%
\item We give two novel algorithms: the first, naive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  Both
  approaches provide sound synthesis of state-feedback systems and consider
  the various sources of imprecision in the implementation of the control
  algorithm and in the modeling of plant.
%
\item We develop a model for different sources of quantization errors and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.
%
\end{enumerate}

@@

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable and under full state
information (i.e., we have access to all the model variables):
%
\begin{align}
\label{eq:ode}
\dot{x}(t) = Ax(t)+ B u(t), \quad x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}, 
\end{align}
%
where $t \in \mathbb R_0^+$, where $A$ and $B$ are matrices that fully
specify the continuous plant, and with initial states set as $x(0)$.  While
we ideally would like to work on the continuous-time plant, in this work
Eq.~\eqref{eq:ode} is soundly discretized in time (as shown in
Appendix~\ref{sec:appendix:LTIbackground}) into
%
%\textcolor{red}{What is subscript ``d''?}
%
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k. 
\end{align} 
%
where $k \in \mathbb N$ and with initial state $x_{0}=x(0)$.  $A_d$ and
$B_d$ denote the matrices that describe the discretized plant dynamics and
$A$ and $B$ the continuous plant dynamics.  Later, we will address the
issue of variable quantization, as introduced by the ADC/DAC conversion
blocks (Fig.~\ref{fig:digitalsystem}).

@@

Models \eqref{eq:ode} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t), u_k$, respectively. 
Feedback architectures can be employed to manipulate the properties and behaviors of the continuous process (the plant).   
%reduce the effect of parametric uncertainties and attenuate the effects of noise and disturbance. 
We are interested in the synthesis of digital feedback control algorithm, 
as implemented on modern Field-Programmable Gate Arrays or Digital Signal Processors. 
The most basic feedback architecture is the state feedback, 
where the control action $u_k$ (notice we work with the discretized signal) is computed by: 
%
\begin{equation}
\label{eq:controlaction}
u_k = r_{k} - K x_k. 
\end{equation}
%
Here $K \in \mathbb{R}^{m \times n}$ is a state-feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital).   
%
The closed-loop model then takes the form 
\begin{align}
\label{eq:closedloopss}
x_{k+1} = ( A_d - B_d K ) x_k + B_d r_k.
\end{align}
%
The gain matrix $K$ can be set so that the closed-loop discrete dynamics are
shaped as desired, for instance according to a specific stability goal or
around a specific dynamical behavior \cite{astrom1997computer}.  As argued
later in this work, we will target more complex objectives, such as
quantitative safety requirements, which are not typical in the digital
control literature.  Further, we will embrace the digital nature of the
controller, which manipulates quantities with finite precision.

>>>>

Linear Time Invariant (LTI) models represent a broad class of dynamical
systems with significant impact in numerous application areas as in the life
sciences, robotics, and engineering~\cite{astrom1997computer,Franklin15}. 
Synthesis of controllers for LTI models is broadly investigated, however the
use of digital control architectures adds new challenges due to the effects
of finite-precision arithmetic, time discretization, and quantization noise,
which are typically introduced by Analogue-to-Digital (ADC) and
Digital-to-Analogue (DAC) conversion.  While research on digital control is
well developed~\cite{astrom1997computer}, automated and sound control
synthesis is challenging, particularly when the synthesis objective goes
beyond classical stability.  There are recent methods for verifying more
general reachability properties of a given controller~\cite{FLD+11,Fre05}. 
However, these methods have not been generalized to control synthesis. 
We~remark that a synthesis algorithm that guarantees stability does not
ensure safety: the system might transitively visit an unsafe state that
results in an unrecoverable failure.

@@

We give two alternative approaches for synthesizing digital controllers for
state-space physical plants, both based on CounterExample Guided Inductive
Synthesis (CEGIS)~\cite{sketch}.  We prove their soundness by quantifying
errors caused by digitalization and quantization effects that arise when
time/value-discrete digital controller interacts with the continuous plant.

@@

We provide experimental results showing that both approaches are able to
efficiently synthesize stable and safe controllers for a set of intricate
physical plant models taken from the digital control literature: the median
run-time for our benchmark set is $7.9$ seconds, and most controllers can be
synthesized in less than $17.2$ seconds.  We further show that, in a direct
comparison, the abstraction-based approach (i.e., second approach) lowers the median
run-time of our benchmarks by a factor of seven over the first approach
based on the unfolding of the dynamics.

@@

\begin{enumerate}
%
\item  We compute state-feedback controllers that are provably stable and in
  addition guaranteee a given safety property.  Existing methods for
  controller synthesis rely on transfer function representations, which are
  insufficient to prove safety requirements.
%
\item We give two novel algorithms: the first, naive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  Both
  approaches provide sound synthesis of state-feedback systems and consider
  the various sources of imprecision in the implementation of the control
  algorithm and in the modeling of the plant.
%
\item We develop a model for different sources of quantization errors and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.
%
\end{enumerate}

@@

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable and under full state
information (i.e., we have access to all the model variables):
%
\begin{align}
\label{eq:ode}
\dot{x}(t) = Ax(t)+ B u(t), \quad x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}, 
\end{align}
%
where $t \in \mathbb R_0^+$, where $A$ and $B$ are matrices that fully
specify the continuous plant, and with initial states set as $x(0)$.  While
we ideally would like to work on the continuous-time plant, in this work
Eq.~\eqref{eq:ode} is soundly discretized in time (as shown in
Appendix~\ref{sec:appendix:LTIbackground}) into
%
%\textcolor{red}{What is subscript ``d''?}
%
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k. 
\end{align} 
%
where $k \in \mathbb N$ and $x_{0}=x(0)$ is the initial state.  $A_d$ and
$B_d$ denote the matrices that describe the discretized plant dynamics, whereas
$A$ and $B$ the continuous plant dynamics.  Later, we will address the
issue of variable quantization, as introduced by the ADC/DAC conversion
blocks (Fig.~\ref{fig:digitalsystem}).

@@

Models \eqref{eq:ode} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t)$ and  $u_k$, respectively. 
Feedback architectures can be employed to manipulate the properties and behaviors of the continuous process (the plant).   
%reduce the effect of parametric uncertainties and attenuate the effects of noise and disturbance. 
We are interested in the synthesis of digital feedback control algorithms, 
as implemented on modern Field-Programmable Gate Arrays or Digital Signal Processors. 
The most basic feedback architecture is the state feedback, 
where the control action $u_k$ (notice we work with the discretized signal) is computed by: 
%
\begin{equation}
\label{eq:controlaction}
u_k = r_{k} - K x_k. 
\end{equation}
%
Here $K \in \mathbb{R}^{m \times n}$ is a state-feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital).   
%
The closed-loop model then takes the form 
\begin{align}
\label{eq:closedloopss}
x_{k+1} = ( A_d - B_d K ) x_k + B_d r_k.
\end{align}
%
The gain matrix $K$ can be set so that the closed-loop discrete dynamics are
shaped as desired, for instance according to a specific stability goal or
around a specific dynamical behavior \cite{astrom1997computer}.  As argued
later in this work, we will target more complex objectives, such as
quantitative safety requirements, which are not typical in the digital
control literature.  Further, we will embrace the digital nature of the
controller, which manipulates quantities with finite precision.
